---
title: "Radiation and Optics"
subtitle: "From BTV1 Intro translated to Quarto"
bibliography: handbook.bib
format: html
editor: source
---
```{r, include=FALSE}
library(ggspectra)
library(photobiologyWavebands)
library(photobiologySun)
library(photobiologyPlants)
```

## Wavelength and Colour

Radiation can be described as waves, with wavelength or frequency as its properties, or based on quantum physics as photons of specific energy. This "double personality" of light is important as each viewpoint provides a simpler explanation for different phenomena, such as refraction an reflection at boundaries between materials and absorption of radiation by pigments and other molecules. From the first perspective, ultraviolet (UV), visible (VIS), and infrared (IR) radiation are electromagnetic waves and are described by the Maxwell's equations.[^1] From the second perspective, Plank's law describes the energy of a quantum as a function of wavelength.

[^1]: These equations are a system of four partial differential equations describing classical electromagnetism.

The wavelength ranges of visible radiation and their usual names, based on human vision, are listed in Table [\[tab:Waverange\]](#tab:Waverange){reference-type="ref" reference="tab:Waverange"}. The colour ranges indicated in Table [\[tab:Waverange\]](#tab:Waverange){reference-type="ref" reference="tab:Waverange"} are an approximation, and we follow the ISO and CIE standards when applicable. Alternative definitions are in use, some of them applicable to plants. The electromagnetic spectrum is continuous with no clear boundaries between one colour and the next. Colours are the result of the interaction of different wavelengths of radiation with the photoreceptor pigments in the eye and the processing of the stimuli by the brain.

```{r, echo=FALSE, warning=FALSE}
ggplot(normalise(sun_reference.mspct$ASTM.G173.global)) +
  geom_line() +
  stat_wb_box(w.band = VIS_bands("ISO"), ypos.fixed = 2, colour = "grey80", box.height = 0.2) +
  stat_wb_label(w.band = VIS_bands("ISO"), ypos.fixed = 2, colour = "white", angle = 90, size = 2) +
  stat_wb_box(w.band = c(UV_bands("ISO"), IR_bands("ISO")), ypos.fixed = 1.8, colour = "grey80", box.height = 0.2) +
  stat_wb_label(w.band = c(UV_bands("ISO"), IR_bands("ISO")), ypos.fixed = 1.8, colour = "white", angle = 90, size = 2) +
  stat_wb_box(w.band = c(UV_bands("ISO"), IR_bands("CIE")), ypos.fixed = 1.6, colour = "grey80", box.height = 0.2) +
  stat_wb_label(w.band = c(UV_bands("ISO"), IR_bands("CIE")), ypos.fixed = 1.6, colour = "white", angle = 90, size = 2) +
  stat_wb_box(w.band = Plant_bands("sensory"), ypos.fixed = 1.4, colour = "grey80", box.height = 0.2) +
  stat_wb_label(w.band = Plant_bands("sensory"), ypos.fixed = 1.4, colour = "white", angle = 90, size = 2) +
  stat_wb_box(w.band = Plant_bands("CIE"), ypos.fixed = 1.2, colour = "grey80", box.height = 0.2) +
  stat_wb_label(w.band = Plant_bands("CIE"), ypos.fixed = 1.2, colour = "white", angle = 90, size = 2) +
  scale_fill_identity() +
  scale_y_s.e.irrad_continuous(breaks = c(0, 0.5, 1), normalized = 531) +
  scale_x_wl_continuous(limits = c(280, 900)) +
  theme_minimal()
```

```{r, echo=FALSE, warning=FALSE}
ggplot(normalise(trim_spct(sun_reference.mspct$ASTM.G173.global, range = c(100, NA), fill = 0))) +
  geom_line() +
  stat_wb_box(w.band = VIS("ISO"), ypos.fixed = 2, colour = "grey80", box.height = 0.2) +
  stat_wb_label(w.band = VIS("ISO"), ypos.fixed = 2, colour = "white", angle = 90, size = 2) +
  stat_wb_box(w.band = c(UV_bands("ISO"), IR_bands("ISO")), ypos.fixed = 1.8, colour = "grey80", box.height = 0.2) +
  stat_wb_label(w.band = c(UV_bands("ISO"), IR_bands("ISO")), ypos.fixed = 1.8, colour = "white", angle = 90, size = 2) +
  stat_wb_box(w.band = c(UV_bands("ISO"), IR_bands("CIE")), ypos.fixed = 1.6, colour = "grey80", box.height = 0.2) +
  stat_wb_label(w.band = c(UV_bands("ISO"), IR_bands("CIE")), ypos.fixed = 1.6, colour = "white", angle = 90, size = 2) +
  stat_wb_box(w.band = PAR(), ypos.fixed = 1.4, colour = "grey80", box.height = 0.2) +
  stat_wb_label(w.band = PAR(), ypos.fixed = 1.4, colour = "white", angle = 90, size = 2) +
  stat_wb_box(w.band = PAR("ePAR"), ypos.fixed = 1.2, colour = "grey80", box.height = 0.2) +
  stat_wb_label(w.band = PAR("ePAR"), ypos.fixed = 1.2, colour = "white", angle = 90, size = 2) +
  scale_fill_identity() +
  scale_y_s.e.irrad_continuous(breaks = c(0, 0.5, 1), normalized = 531) +
  scale_x_wl_continuous(limits = c(NA, 3000)) +
  theme_minimal()
```

Different regions or *wavebands* of UV and IR radiation, are also listed in Table [\[tab:Waverange\]](#tab:Waverange){reference-type="ref" reference="tab:Waverange"}. It is important to keep in mind that the ISO and CIE definitions of the VIS and UV wavebands overlap. Thus, it is incorrect to sum VIS and UV irradiances, while the definitions of photosynthetically active radiation (PAR) and UV do not overlap. Especially in the IR region the subdivision is somewhat arbitrary and the boundaries used in the literature vary, with differences even between ISO and CIE standards.

The energy of a quantum of radiation in a vacuum, depends on the wavelength, $\lambda$, or frequency[^2], $\nu$, 

$$q = h \cdot \nu = h \cdot \frac{c}{\lambda}$$ 

with the Planck constant $h=6.626\times 10^{-34}\,\mathrm{J\,s}$ and speed of light in vacuum $c=2.998\times 10^{8}\,\mathrm{m}\,\mathrm{s}^{-1}$. When dealing with numbers of photons, the equation above can be extended by using Avogadro's number $N_\mathrm{A}=6.022\times 10^{23}$ mol$^{-1}$. Thus, the energy of one mole of photons, $q^\prime$, is 

$$q^\prime = h^\prime \cdot \nu = h^\prime \cdot \frac{c}{\lambda}$$ 
with $h'=h\cdot N_\mathrm{A}=3.990\times 10^{-10}\,\mathrm{J\,s\,mol}^{-1}$. 

The two equations above, are valid for all kinds of electromagnetic waves.
 
:::callout-note
# examples

**Example 1:** red light at 600 nm has about 200 kJ mol$^{-1}$, therefore, 1 $\mu$mol photons has 0.2 J. **Example 2:** radiation at 300 nm has about 400  kJ mol$^{-1}$, therefore, 1 $\mu$mol photons has 0.4 J.
:::

[^2]: Wavelength and frequency are related to each other by the speed of light, according to $\nu = c / \lambda$ where $c$ is speed of light in vacuum. Consequently there are two equivalent formulations for the equations.


## Interactions of Light and Matter

The processes responsible for the variation of the radiance $L(\lambda,\theta,\phi)$ as the radiation beam travels through any kind of material, are primarily absorption $a$ and scattering $b$, which are called inherent optical properties, because they depend only on the characteristics of the material itself and are independent of the light field. Radiance is added to the directly transmitted beam, coming from different directions, due to elastic scattering, by which a photon changes direction but not wavelength or energy level. An example of this is Raleigh scattering in very small particles, which causes the scattering of light in a rainbow. A further gain of radiance into the direct path is due to inelastic processes like fluorescence, where a photon is absorbed by the material and reemitted as a photon with a longer wavelength and lower energy level, and Raman scattering. The elastic and inelastic scattered radiance is denoted as $L^E$ and $L^I$, respectively. Internal sources of radiances, $L^S$, like bioluminescence of biological organisms or cells contribute also to the detected radiance. The path of the radiance through a thin horizontal layer with thickness d$z=z_1-z_0$ is shown schematically in Figure [\[fig_radtrans\]](#fig_radtrans){reference-type="ref" reference="fig_radtrans"}.

::: figure*
::: center
::: picture
(10,6) (0,1)(1,0)10 (0,5)(1,0)10 (7,1)(1,-1)1 (2,6)(1,-1)1 (3,5)(1,-1)4 (4,4) (6,2) (4,4)(1,1)0.53 (4,4)(1,0)0.75 (4,4)(-1,0)0.75 (4,4)(-1,-1)0.53 (5.1,2)(1,0)0.75 (6.6,2.6)(-1,-1)0.53 (6,2)(1,-1)0.75 (0,0.7)$z_1$(0,4.7)$z_0$ (3.25,5.7)Incoming radiance (3.25,5.2)$L(\theta,\phi)$ (5,4.5)Loss by absorption $a$ (5,4)and scattering $b$ (7,3)Gain by elastic (7,2.5)scattered radiance (7,2)$L^E$ into the path (0.25,2.25)Gain by inelastic scattered (0.25,1.75)radiance $L^I$ and internal (0.25,1.25)sources $L^S$ into the path (4.2,0.6)Outgoing radiance (4.2,0.1)$L(\theta,\phi)$
:::
:::
:::

## Angle of Incidence

Most frequently we describe the flux of radiation on a plane, usually, but not always a horizontal plane. This called *irradiance* and expressed per unit area and unit time. The area used as a reference is that of the receiving plane. This plane is located in three-dimensional (3D) space and when considering a point source of radiation the angle between the light beam and the receiving plane or surface affects the flux of radiation.

The sun, can be assumed to be a point source (the solar disk is $\approx 1 ^\circ$ in diameter when seen from Earth), and changes in distance small relative to the total Sun to Earth distance, ignored. Thus, we have to consider only the angle of incidence @fig-cosine. For a given light source the maximum irradiance is that received on a plane normal to the light beam. As the angle between the beam and the surface becomes shallower, the irradiance measured on the plane decreases, reaching zero, when the beam is parallel to the surface. The relationship between the normal area and that at shallower angle is given by the cosine of the angle. When considering the sun, if the receiving plane is horizontal, the angle of interest is the solar elevation (1 - zenith angle). If the plane is not horizontal, the angle between the plane and the light beam depends both the the zenith angle and the azimuth angle of the sun's position, and the effective angle needs to be computed in three dimensions.

Obviously, the discussion above ignored scattered radiation from the sky. One approach to the simulation of the light field when multiple sources or sources with a large emitting area are present, is to use ray tracing, which is essentially a divide and conquer approach.

When we considered a point source like an LED or light small light bulb at a relatively short distance from the receiving plane, distance cannot be ignored, as the illuminated area increases four times for each doubling in the distance. Thus, each doubling in distance decreases irradiance by 75%.

In the case of light sources of a large surface or light beams focused by a lens or reflector, the rate of increase in area is smaller than for a point source. The extreme case is a highly collimated laser beam, for which the illuminated area increases very slowly with increasing distance.

::: callout-tip
Using a lens or reflector to make a light beam narrower makes it possible to increase irradiance by concentrating the light energy onto a smaller illuminated area, and this effect can be quantitatively large. Alternatively, the narrower beam makes it possible to move the light source away from the illuminated object maintaining the same irradiance. This can provide additional space for other equipment, or when using selective reflectors that transmit infrared radiation and only reflect VIS and/or UV radiation reduce the heat load on the illuminated object.
:::

In a perfectly collimated beam, the photons follow parallel trajectories. In a beam from a ideal point source the trajectories of the photons are equally probably in all directions (away from the centre of an imaginary sphere). Most real light sources behave somewhere inbetween these extremes.

::: callout-note
# Angle of Incidence in 3D

While in above, we had the sensor or observer at the centre, and the light source located on an imaginary sphere, here, instead, we locate the point light source at the centre of the sphere. So, when a beam or the radiation passing into a space or sphere is analysed, two important parameters are necessary: the distance to the source and the position of the measurement plane---i.e., if the receiving surface is perpendicular to the beam or at a shallower angle. The geometry is illustrated in Figure [\[fig_space\]](#fig_space){reference-type="ref" reference="fig_space"}. The radiation is received at distance $r$ by a surface of area d$A$, tilted by an angle $\alpha$ to the unit sphere's surface element, a solid angle, d$\Omega$, which is a two-dimensional angle in a space. The relation between d$A$ and d$\Omega$ in spherical coordinates is geometrically explained in Figure [\[fig_space\]](#fig_space){reference-type="ref" reference="fig_space"}.

::: figure*
![image](figures/sphere_highres.pdf){width="0.8\\myfigwidth"}
:::

The solid angle is calculated from the zenith angle $\theta$ and azimuth angle $\phi$, which denote the direction of the radiation beam $${\rm d}\Omega = {\rm d}\theta\cdot\sin\theta{\rm d}\phi \label{equ_dsolidangle}$$ The area of the receiving surface is calculated by a combination of the solid angle of the beam, the distance $r$ from the radiation source and the angle $\alpha$ of the tilt: $${\rm d}A  =  \frac{r{\rm d}\theta}{\cos\alpha}\cdot r\sin\theta{\rm d}\phi$$ which can be rearranged to $$\Rightarrow {\rm d}A  =  \frac{r^2}{\cos\alpha}\quad{\rm d}\Omega$$ Thus, the solid angle is given by $$\Omega = \int_A \frac{{\rm d}A\cdot\cos\alpha}{r^2} \label{equ_solidangle}$$ The unit of the solid angle is a steradian (sr). The solid angle of an entire sphere is calculated by integration of equation ([\[equ_dsolidangle\]](#equ_dsolidangle){reference-type="ref" reference="equ_dsolidangle"}) over the zenith ($\theta$) and azimuth ($\phi$) angles, $0\le\theta\le\pi(180^\circ)$ and $0\le\phi\le2\pi(360^\circ)$, and is $4\pi$ sr.
:::

## Radiative Transfer

Putting all this together, the radiative transfer equation is $$\cos\theta\,\frac{{\rm d}L}{{\rm d}z} = -(a+b)\cdot L + L^E + L^I + L^S \label{equ_rte}$$ The dependencies of $L$ on $\lambda$, $\theta$, and $\phi$ are omitted here for brevity. No exact analytical solution to the radiative transfer equation exists, hence it is necessary either to use numerical models or to make approximations and find an analytical parameterisation. A numerical model is for example the Monte Carlo method. The parameters of the light field can be simulated by modelling the paths of photons. For an infinite number of photons the light field parameters reach their exact values asymptotically. The advantage of the Monte Carlo method is a relatively simple structure of the program, and it simulates nature in a straightforward way, but its disadvantage is the time-consuming computation involved. Details of the Monte Carlo method are explained for example by @prahl89, @Wang1995[^3], or @moble94.

[^3]: Their program is available from the website of [Oregon Medical Laser Center](http://omlc.ogi.edu/software/mc/) at <http://omlc.ogi.edu/software/mc/>

The other way to solve the radiative transfer equation is through the development of analytical parameterisations by making approximations for all the quantities needed. In this case, the result is not exact, but it has the advantage of fast computing and the analytical equations can be inverted just as fast. This leads to the idealised case of a source-free ($L^S=0$) and non-scattering media, i.e., $b=0$ and therefore $L^E=L^I=0$. Then, equation [\[equ_rte\]](#equ_rte){reference-type="ref" reference="equ_rte"} can be integrated easily and yields $$%L(z_1) = L(z_0)\cdot\exp\left\{-\frac{a\cdot(z_1-z_0)}{\cos\theta}\right\} \label{eq:Beer-Lambert}
L(z_1) = L(z_0)\cdot \mathrm{e}^{-\frac{a\cdot(z_1-z_0)}{\cos\theta}} \label{eq:Beer-Lambert}$$ The boundary value $L(z_0)$ is presumed known. This result is known as Beer's law (or Lambert's law, Bouguer's law, Beer-Lambert law), denotes any instance of exponential attenuation of light and is exact only for purely absorbing media---i.e., media that do not scatter radiation. It is of direct application in analytical chemistry, as it describes the direct proportionality of absorbance ($A$) to the concentration of a coloured solute in a transparent solvent.

## Quantities

::: table*
::: center
|                                                                 Symbol                                                                 |         Unit         | Description                                                                                                                                          |
|:------------------------:|:---------------:|:-----------------------------|
|                                              $\Phi=\frac{\partial \quantum}{\partial t}$                                               |    W = J s$^{-1}$    | Radiant flux: absorbed or emitted energy per time interval                                                                                           |
|                                            $\exposure=\frac{\partial \quantum}{\partial A}$                                            |      J m$^{-2}$      | Exposure: energy towards a surface area. (In plant research this is called usually *dose* (), while in Physics *dose* refers to absorbed radiation.) |
|                                                $\irr=\frac{\partial \Phi}{\partial A}$                                                 |      W m$^{-2}$      | Irradiance: flux or radiation towards a surface area, radiant flux density                                                                           |
|                                           $\intensity=\frac{\partial \Phi}{\partial\Omega}$                                            |     W sr$^{-1}$      | Radiant intensity: emitted radiant flux of a surface area per solid angle                                                                            |
|                                               $\emitt=\frac{\partial \Phi}{\partial A}$                                                |      W m$^{-2}$      | Emittance: emitted radiant flux per surface area                                                                                                     |
| $\radiance=\frac{\partial^2 \Phi}{\partial \Omega (\partial A\cdot \cos\alpha)}=\frac{\partial \intensity}{\partial A\cdot\cos\alpha}$ | W m$^{-2}$ sr$^{-1}$ | Radiance: emitted radiant flux per solid angle and surface area depending on the angle between radiant flux and surface perpendicular                |
:::
:::

Different physical quantities are used to describe the "amount of radiation" and their definitions and abbreviations are listed in Table [\[tab:Phys:Quants\]](#tab:Phys:Quants){reference-type="ref" reference="tab:Phys:Quants"}. Taking into account Equation [\[equ_solidangle\]](#equ_solidangle){reference-type="ref" reference="equ_solidangle"} and assuming a homogenous flux, the important correlation between irradiance $\irr$ and intensity $\intensity$ is $$E = \frac{\intensity \cdot \cos\alpha}{r^2} \label{equ_r2law}$$ The irradiance decreases by the square of the distance to the source and depends on the tilt of the detecting surface area. This is valid only for point sources. For outdoor measurements the sun can be assumed to be a point source. For artificial light sources simple LEDs (light-emitting diodes) without optics on top are also effectively point sources. However, LEDs with optics---and other artificial light sources with optics or reflectors designed to give a more focused dispersal of the light---deviate to various extents from the rule of a decrease of irradiance proportional to the square of the distance from the light source.

Besides the physical quantities used for all electromagnetic radiation, there are also equivalent quantities to describe visible radiation, so called photometric quantities. The human eye as a detector led to these photometric units, and they are commonly used by lamp manufacturers to describe their artificial light sources. See Box [\[box:photometric:quantities\]](#box:photometric:quantities){reference-type="ref" reference="box:photometric:quantities"} on page for a short description of these quantities and units.

::: example*
:::

::: example*
::: shaded
When we are interested in photochemical reactions, the most relevant radiation quantities are those expressed in photons. The reason for this is that, as discussed in section [1.2](#sec:photobio:principles){reference-type="ref" reference="sec:photobio:principles"} on page , molecules are excited by the absorption of certain fixed amounts of energy or quanta. The surplus energy "decays" by non-photochemical processes. When studying photosynthesis, where many photons of different wavelengths are simultaneously important, we normally use photon irradiance to describe amount of . The name photosynthetic photon flux density, or , is also frequently used when referring to photon irradiance. When dealing with energy balance of an object instead of photochemistry, we use (energy) irradiance. In meteorology both and visible radiation, are quantified using energy-based quantities. When dealing with photochemistry as in responses mediated by UVR8, an photoreceptor, the use of quantum quantities is preferred. According to the physical energetic quantities in the table [\[tab:Phys:Quants\]](#tab:Phys:Quants){reference-type="ref" reference="tab:Phys:Quants"}, the equivalent photon related quantities are listed in the table below and have the subscript $\mathrm{p}$.

**Table.** Photon quantities of light.

::: center
|                       Symbol                       |       Unit        | Description                                                                                                         |
|:---------------:|:---------------:|:-------------------------------------|
|                 $\Phi_\mathrm{p}$                  |     s$^{-1}$      | Photon flux: number of photons per time interval                                                                    |
| $\pfd=\frac{\partial \Phi_\mathrm{p}}{\partial A}$ | m$^{-2}$ s$^{-1}$ | Photon irradiance: photon flux towards a surface area, photon flux density (sometimes also symbolised by $\irr[p]$) |
|         $\qdose=\int_t\;\pfd\;\mathrm{d}t$         |     m$^{-2}$      | Photon exposure: number of photons towards a surface area during a time interval, photon fluence                    |
:::

These quantities can be also used based on a 'chemical' amount of moles by dividing the quantities by Avogadro's number $N_A=6.022\times 10^{23}$ mol$^{-1}$. To determine a quantity in terms of photons, an energetic quantity has to be weighted by the number of photons, i.e. divided by the energy of a single photon at each wavelength as defined in equation [\[equ_energy\]](#equ_energy){reference-type="ref" reference="equ_energy"}. This yields for example $$\Phi_\mathrm{p} = \frac{\lambda}{h\;c}\cdot\frac{\partial\quantum}{\partial t}\hspace{1cm}\mathrm{and}\hspace{1cm}\pfd(\lambda) = \frac{\lambda}{h\;c}\cdot E(\lambda)$$ When dealing with bands of wavelengths, for example an integrated value like from 400 to 700 nm, it is necessary to repeat these calculations at each wavelength and then integrate over the wavelengths. For example, the photon irradiance or in moles of photons is obtained by $$\PPFD = \frac{1}{N_\mathrm{A}}\int_{400\;\mathrm{nm}}^{700\,\mathrm{nm}}\frac{\lambda}{hc}\;E(\lambda)\;\mathrm{d}\lambda$$ For integrated values of or radiation the calculation is done analogously by integrating from 280 to 315 nm or 315 to 400 nm, respectively.

If we have measured (energy) irradiance, and want to convert this value to photon irradiance, the exact conversion will be possible only if we have information about the spectral composition of the measured radiation. Conversion factors at different wavelengths are given in the table below. For , 1 of "average daylight" is approximately 4.6 . This is exact only if the radiation is equal from 400 to 700 nm, because the factor is the value at the central wavelength at 550 nm. Further details are discussed in section [\[sec:basic:concepts\]](#sec:basic:concepts){reference-type="ref" reference="sec:basic:concepts"} on page .

**Table.** Conversion factors of photon and energy quantities at different wavelengths.

::: center
|     |  to  | $\lambda$ (nm) |
|-----|:----:|:--------------:|
|     | 2.34 |      280       |
|     | 2.49 |      298       |
|     | 2.63 |      315       |
|     | 2.99 |      358       |
|     | 3.34 |      400       |
|     | 4.60 |      550       |
|     | 5.85 |      700       |
:::
:::
:::

There are, in principle, two possible approaches to measuring radiation. The first is to observe light from one specific direction or viewing angle, which is the radiance $\radiance$. The second is to use a detector, which senses radiation from more than one direction and measures the so-called irradiance $\irr$ of the entire sphere or hemisphere. The correlation between irradiance $\irr$ and radiance $\radiance$ of the wavelength $\lambda$ is given by integrating over all directions of incoming photons. $$\begin{aligned}
\irr[0](\lambda) & = \int_\Omega \radiance(\lambda,\Omega) {\rm d}\Omega \label{equ_E0} \\
\irr(\lambda) & = \int_\Omega \radiance(\lambda,\Omega) |\cos\alpha| {\rm d}\Omega \label{equ_E}
%\end{eqnarray}
\end{aligned}$$ Depending on the shape of a detector (which may be either planar or spherical) the irradiance is called (plane) irradiance $\irr$ or fluence rate (also called scalar irradiance) $\flrat$. A planar sensor detects incoming photons depending on the incident angle and a spherical sensor detects all photons equally weighted for all directions. See section [\[sec:basic:concepts\]](#sec:basic:concepts){reference-type="ref" reference="sec:basic:concepts"} on page for a more detailed discussion.

Here we have discussed the properties of light based on energy quantities. In photobiology there are good reasons to quantify radiation based on photons. See Box [\[box:quantum:quantities\]](#box:quantum:quantities){reference-type="ref" reference="box:quantum:quantities"} on page , and section [\[sec:basic:concepts\]](#sec:basic:concepts){reference-type="ref" reference="sec:basic:concepts"} on page .
